Certainly, here's the revised tutorial using the `tsibble` package instead of `sibble` for time series analysis and forecasting with the `fable` package.

```{r setup, include=FALSE}
library(tsibble)
library(fable)
library(ggplot2)
```

# Tutorial: Time Series Analysis and Forecasting with tsibble and fable

## Introduction

In this tutorial, we will explore the use of the `tsibble` and `fable` packages in R to analyze time series data and perform forecasting. We will use the famous AirPassengers dataset, which contains monthly airline passenger counts from 1949 to 1960.

### Task 1: Analyzing the Data (1949-1955)

**Objective:** Fit a regression model with seasonal and trend decomposition to the data from 1949 to 1955.

**Instructions:** 1. Load the AirPassengers dataset. 2. Subset the data to include only the period from 1949 to 1955. 3. Create a tsibble object from the subset data. 4. Fit a regression model with seasonal and trend decomposition using `fable`. 5. Visualize the decomposition components.

```{r task1}
# Load the AirPassengers dataset
data("AirPassengers")


# Create a tsibble object
passengers_tbl <- AirPassengers %>%
  as_tsibble(key = "Month")

passengers_data <- AirPassengers[1:84,]

# Fit a regression model with seasonal and trend decomposition
decomposition_model <- passengers_tbl %>%
  model(
    decomposition = ETS(value ~ trend() + season())
  )

# Plot decomposition components
decomposition_model %>%
  components() %>%
  autoplot()
```

**Hints:** - Use the `as_tsibble()` function to convert your data to a tsibble object. - When fitting the model, use the `ETS` method with `trend()` and `season()` components.

**Example Solution:** The code provided above will create a decomposition plot showing the trend and seasonal components of the time series data.

### Task 2: Making Forecasts (1956-1960) and Comparison

**Objective:** Generate forecasts for each month in the years 1956-1960 and compare them with the actual data.

**Instructions:** 1. Subset the data to include only the period from 1956 to 1960. 2. Create a tsibble object from the subset data. 3. Generate forecasts using the previously fitted model. 4. Plot the forecasts and compare them with the actual data from the AirPassengers dataset.

```{r task2}
# Subset data from 1956 to 1960
forecast_data <- subset(AirPassengers, subset = year(Month) >= 1956 & year(Month) <= 1960)

# Create a tsibble object for the forecast period
forecast_tbl <- forecast_data %>%
  as_tsibble(key = "Month")

# Generate forecasts
forecasts <- decomposition_model %>%
  forecast(new_data = forecast_tbl)

# Plot forecasts and compare with actual data
compare_data <- cbind(actual = forecast_tbl$Passengers, forecasted = as.numeric(forecasts$forecast))

ggplot(compare_data, aes(x = Month)) +
  geom_line(aes(y = actual, color = "Actual")) +
  geom_line(aes(y = forecasted, color = "Forecasted")) +
  labs(title = "Actual vs. Forecasted AirPassengers Data (1956-1960)")
```

**Hints:** - Use the `forecast()` function with the previously fitted decomposition model to generate forecasts.

**Example Solution:** The code provided above will create a plot that compares the actual and forecasted data for the years 1956-1960.

This concludes our tutorial on time series analysis and forecasting using the `tsibble` and `fable` packages. You've learned how to decompose time series data and make forecasts. Feel free to explore further and customize your analysis as needed.